This code is taken from https://www.kaggle.com/code/erikbruin/titanic-2nd-degree-families-and-majority-voting/report
```{r}
knitr::opts_chunk$set(echo = TRUE)
```

First we load all required packages
```{r}
library(Hmisc)
library(knitr)
library(ggplot2)
library(dplyr)
library(caret)
library(randomForest)
library(gridExtra)
library(ROCR)
library(corrplot)
```
and then load the dataset into training and testing data

```{r}
train <- read.csv("C:/Users/Danie/OneDrive/Documents/_Uni/Maths/Year 4/Data Science Toolkit/DST_Assessment_0/Report/01-Data.csv", stringsAsFactors = F, na.strings = c("NA", ""))
test <- read.csv("C:/Users/Danie/OneDrive/Documents/_Uni/Maths/Year 4/Data Science Toolkit/DST_Assessment_0/Report/01-Data.csv", stringsAsFactors = F, na.strings = c("NA", ""))
```

We then set the survival column in test to N/A's and combine the two

```{r}
test$Survived <- NA
all <- rbind(train, test)

all$Sex <- as.factor(all$Sex)
all$Survived <- as.factor(all$Survived)
all$Pclass <- as.ordered(all$Pclass)
```

We can then start visualizing the amount of deaths and who died

```{r}
ggplot(all[!is.na(all$Survived),], aes(x = Survived, fill = Survived)) +
  geom_bar(stat='count') +
  labs(x = 'How many people died and survived on the Titanic?') +
        geom_label(stat='count',aes(label=..count..), size=7) +
        theme_grey(base_size = 18)
```

And we can do further analysis on how class affects your chances of survival

```{r}
p3 <- ggplot(all, aes(x = Pclass, fill = Pclass)) +
  geom_bar(stat='count', position='dodge') +
  labs(x = 'Pclass, All data') + geom_label(stat='count', aes(label=..count..)) +
   theme(legend.position="none") + theme_grey()     
p4 <- ggplot(all[!is.na(all$Survived),], aes(x = Pclass, fill = Survived)) +
  geom_bar(stat='count', position='dodge') + labs(x = 'Training data only') +
        theme(legend.position="none") + theme_grey()
p5 <- ggplot(all[!is.na(all$Survived),], aes(x = Pclass, fill = Survived)) +
  geom_bar(stat='count', position='stack') +
  labs(x = 'Training data only', y= "Count") + facet_grid(.~Sex) +
        theme(legend.position="none") + theme_grey()
p6 <- ggplot(all[!is.na(all$Survived),], aes(x = Pclass, fill = Survived)) +
  geom_bar(stat='count', position='fill') +
  labs(x = 'Training data only', y= "Percent") + facet_grid(.~Sex) +
        theme(legend.position="none") + theme_grey()

grid.arrange(p3, p4, p5, p6, ncol=2)
```
And even the realtion between the size of family and chances of survival

```{r}
#creating family size variable (Fsize)
all$Fsize <- all$SibSp+all$Parch +1
ggplot(all[!is.na(all$Survived),], aes(x = Fsize, fill = Survived)) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') + theme_grey()
```

