---
title: "Work"
author: "Xin Guan"
date: "04/10/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages

```{r}
library(readr) #For read_csv
library(dplyr)
library(fs) # for cross-platform directories (path_wd)
library(knitr) # For kable 
library(gplots)
library(plotrix)
```

Import Data

```{r}
data <- read_csv("~/Desktop/DST/DST_Assessment_0/Report/01-Data.csv")
data.f= as.data.frame(data)
#read_csv imported data as tbl which is a subclass of data frame, however it behaves differentlt so I convert it to data.frame format.
```
```{r}
head(data.f)
```

```{r}
summary(data.f)
```

Let's check how many missing values in each column.
```{r}
sapply(data.f, function(x) sum(is.na(x)))
```
Delete unwanted columns

```{r}
data.f<-data.f[,c(2:8,10,12)]
```

Deal with missing values in each column.

```{r}
data.f$Embarked[is.na(data.f$Embarked)]<-mode(data.f$Embarked) 
#Use mode for the missing "Embarked" values.

data.f$Age[is.na(data.f$Age)]<-mean(data.f$Age,na.rm = T)
#Use average of the existing age values for the missing "Age" values.
```

Now we have got rid of all the missing values in our data set.

```{r}
sapply(data.f, function(x) sum(is.na(x)))
```


```{r}
labs=unique(as.character(data.f[,"Embarked"]))
names(labs)=labs

list=lapply(labs,function(x){
  data.f[data.f[,"Embarked"]==x,1:8]
})
#Separate data in to three sub-data by the port of embark.
```

```{r}
data.f %>% count(Embarked)
# Count each numbers of each value in the Embarked variable
```

```{r}
#Plot heatmap withou Age variable
datamean2=t(sapply(list,function(x)colMeans(x[,c(2,5,6,7,8)])))
heatmap.2(datamean2,trace="none",margins=c(7,3))
```

From the heat map, we can see that only the fare depend on the port of embark, the other three numerical variables are independent of the port of embark.


```{r}
data.f %>% count(Sex)
```

```{r}
genders=unique(as.character(data.f[,"Sex"]))
names(genders)=genders
genders_data=lapply(genders,function(x){
  data.f[data.f[,"Sex"]==x,]
})
male_data<-genders_data[1]
female_data<-genders_data[2]
female_data= as.data.frame(female_data)
male_data= as.data.frame(male_data)
# Separate data into two sub data by genders
```

```{r}
fe_sur_rate <- sum(female_data$female.Survived)/dim(female_data)[1]

male_sur_rate <- sum(male_data$male.Survived)/dim(male_data)[1]
```

```{r}
fe_sur_rate
male_sur_rate
```

We see that the survival rate of female is a lot higher than male, this is because that the men on board the ocean liner gave women and children priority access to the lifeboats.

```{r}
histStack(x=data.f$Fare,z=factor(data.f$Survived),col=c("red","blue"),main = "Fare Histogram by Survived",ylim=c(0,800))
legend("topright",legend=c("Survived","Dead"),col=c("blue","red"),lwd=5,cex=0.5) #Add legend to indicates data for Survived and Dead
```

People between 20-40 years old has significantly lower  survival rate. 
```{r}
histStack(x=data.f$Age,z=factor(data.f$Survived),col=c("red","blue"),main = "Age Histogram by Survived",ylim=c(0,400))
legend("topright",legend=c("Survived","Dead"),col=c("blue","red"),lwd=5,cex=0.5)
```

Create a variable "Familysize" which is the sum of variables "SibSp" and "Parch".

```{r}
data.f$Familysize=data.f$SibSp+data.f$Parch
```

```{r}
histStack(x=data.f$Familysize,z=factor(data.f$Survived),col=c("red","blue"),main = "Family size Histogram by Survived")
legend("topright",legend=c("Survived","Dead"),col=c("blue","red"),lwd=5,cex=0.5)
```

If we zoom in a little bit, we can see that the family size of 2 or 3 has significantly higher survival rate than family of other sizes.

```{r}
histStack(x=data.f$Familysize,z=factor(data.f$Survived),col=c("red","blue"),main = "Family size Histogram by Survived",xlim=c(1,10),ylim=c(0,100))
legend("topright",legend=c("Survived","Dead"),col=c("blue","red"),lwd=5,cex=0.5)
```